ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-09-30 rebuttal.tex

    [3] Please add reference: \url{https://www.itv.rwth-aachen.de/index.php?id=13}  Also, the
    latest version is from 2017.

    [7] Further info:  Paper P1-04, ISBN 978-963-12-1257-0

    [8] Please refer to ReSpecTh Kinetics Data Format Specification 2.0 instead.
    %
\end{reviewer}

\begin{response}
    %
    We corrected those references; thank you for pointing out the missing\slash old information.
    %
\end{response}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reviewer \#2:}

\begin{reviewer}
    %
    My main concern with the creation of a publically available database, formatted with ChemKED, is
    that it seems that the creation of the files has to be done manually, i.e. susceptible for
    errors. This is especially true if the experiment comprises a big set of datapoints and/or
    hundreds of measured species. Furthermore, the YAML data serialization format is indeed
    human-readable, but it does not provide an easy summary/overview of the experimental data, such
    as an excel or CSV file would give, and therefore, people might be not willing to put the extra
    effort.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer, insofar as manual creation of the files may lead to errors in the
    encoded data. However, this will be the case wherever manual processing is required, regardless
    of the database format, and could even be the case for tables or figures published in journal
    articles. Thus, we do not feel that manual file creation is an inherent disadvantage of our
    format. Moreover, as we discuss below, we are working with experimental teams to integrate the
    creation of ChemKED files into their workflows to try to mitigate some of these problems.

    On the second point, we agree with the reviewer that CSV or Excel files may give an adequate
    summary or overview of the data, but only if the files are opened with some appropriate
    software, e.g., Microsoft Excel (particularly for Excel-formatted files). Thus, in this sense,
    CSV or Excel files do not offer any advantage over ChemKED files (or any other database files)
    that must also be processed with an appropriate software package (PyKED, in the case of ChemKED
    files) so that a summary can be produced. Furthermore, in the specific case of PyKED/ChemKED,
    it is trivial to produce an Excel or CSV file so that one may view such a summary if desired.
    See also the response to a similar comment below.
    %
\end{response}

### 2017-09-30 rebuttal.tex

\begin{response}
    %
    We agree with the reviewer, insofar as manual creation of the files may lead to errors in the
    encoded data. However, this will be the case wherever manual processing is required, regardless
    of the database format, and could even be the case for tables or figures published in journal
    articles. Thus, we do not feel that manual file creation is an inherent disadvantage of our
    format. Moreover, as we discuss below, we are working with experimental teams to integrate the
    creation of ChemKED files into their workflows to try to mitigate some of these problems.

    On the second point, we agree with the reviewer that CSV or Excel files may give an adequate
    summary or overview of the data, but only if the files are opened with some appropriate
    software, e.g., Microsoft Excel (particularly for Excel-formatted files). Thus, in this sense,
    CSV or Excel files do not offer any advantage over ChemKED files (or any other database files)
    that must also be processed with an appropriate software package (PyKED, in the case of ChemKED
    files) so that a summary can be produced. Furthermore, in the specific case of PyKED/ChemKED,
    it is trivial to produce an Excel or CSV file so that one may view such a summary if desired.
    See also the response to a similar comment below.
    %
\end{response}

\begin{reviewer}
    %
    1.~How are the authors planning on `forcing' the experimentalists to publish their data in this
    new ChemKED file format?
    %
\end{reviewer}

\begin{response}
    %
    While we cannot force anyone to use a particular format, we hope that by defining a format and
    ecosystem with desirable features---including inherent openness and usability---experimentalists
    will begin adopting a community-standard format. In fact, we have already received expressions
    of interest from various experimental groups, and hope this will grow as we continue to improve
    and add features. We also hope that at a certain point, with wide adoption, sharing experimental
    data in a standard format will become the expected norm (e.g., when submitting an article for
    publication) like sharing a Chemkin-format chemical kinetic model.

    Furthermore, we are willing to work (and are currently working) directly with experimentalists
    to provide them easy-to-use conversion scripts from internal data formats. In return, they are
    helping us define what data and information needs to go in the specification for their
    experiment.
    %
\end{response}

### 2017-09-30 rebuttal.tex

\begin{response}
    %
    While we cannot force anyone to use a particular format, we hope that by defining a format and
    ecosystem with desirable features---including inherent openness and usability---experimentalists
    will begin adopting a community-standard format. In fact, we have already received expressions
    of interest from various experimental groups, and hope this will grow as we continue to improve
    and add features. We also hope that at a certain point, with wide adoption, sharing experimental
    data in a standard format will become the expected norm (e.g., when submitting an article for
    publication) like sharing a Chemkin-format chemical kinetic model.

    Furthermore, we are willing to work (and are currently working) directly with experimentalists
    to provide them easy-to-use conversion scripts from internal data formats. In return, they are
    helping us define what data and information needs to go in the specification for their
    experiment.
    %
\end{response}

\begin{reviewer}
    %
    2.~Am I correct to state that ChemKED is a YAML file with a predefined structure?
    %
\end{reviewer}

\begin{response}
    %
    Yes, ChemKED files are files in YAML format, with a prescribed structure and set of contents.
    The required structure and contents are defined by our schema, which is documented in this
    manuscript and in the online documentation.
    %
\end{response}

@@

\end{response}

\subsection*{Abstract and introduction}

\begin{reviewer}
    %
    4. In the abstract, it is mentioned that the Python-based package PyKED creates ChemKED-
    formatted files. However in the article there is no evidence that PyKED aides in creating
    ChemKED-formatted files.
    %
\end{reviewer}

\begin{response}
    %
    The reviewer is correct, and this was a typo in the abstract. We have clarified that PyKED is
    intended as software to validate and work with ChemKED files.

    That said, we do have ongoing work to create a graphical user interface to create ChemKED files,
    as this was requested by several of the experimental teams that have expressed interest in the
    ChemKED format. The work is available publicly at
    \url{https://github.com/pr-omethe-us/ChemKED-gui}.
    %
\end{response}

### 2017-09-30 pyked.tex
Respond to #21
Point out the ongoing GUI work to create ChemKED files, and clarify that
PyKED is for validating and working with CK files

%====================================================================
\begin{abstract} % not to exceed 200 words
Fundamental experimental measurements of quantities such as ignition delay times,
laminar flame speeds, and species profiles (among others) serve important roles in understanding
fuel chemistry and validating chemical kinetic models. However, despite both the
importance and abundance of such information in the literature, the community lacks
a widely adopted standard format for this data. This impedes both sharing and wide
use by the community. Here we introduce a new chemical kinetics
experimental data format, ChemKED, and the related Python-based package for creating
and validating ChemKED-formatted files called PyKED. We also review past and
related efforts, and motivate the need for a new solution. ChemKED currently supports
the representation of autoignition delay time measurements from shock tubes and rapid
compression machines.
ChemKED-formatted files contain all of the information needed to simulate experimental
data points, including the uncertainty of the data. ChemKED is based on the YAML data
serialization language, and is intended as a human- and machine-readable standard for
easy creation and automated use.
Development of ChemKED and PyKED occurs openly on GitHub under the
BSD 3-clause license, and contributions from the community are welcome. Plans for
future development include support for experimental data from laminar flame,
jet stirred reactor, and speciation measurements.
\end{abstract}

>>>>

%====================================================================
\begin{abstract} % not to exceed 200 words
    %
    Fundamental experimental measurements of quantities such as ignition delay times, laminar flame
    speeds, and species profiles (among others) serve important roles in understanding fuel
    chemistry and validating chemical kinetic models. However, despite both the importance and
    abundance of such information in the literature, the community lacks a widely adopted standard
    format for this data. This impedes both sharing and wide use by the community. Here we introduce
    a new chemical kinetics experimental data format, ChemKED, and the related Python-based package
    for validating and working with ChemKED-formatted files called PyKED. We also review past and
    related efforts, and motivate the need for a new solution. ChemKED currently supports the
    representation of autoignition delay time measurements from shock tubes and rapid compression
    machines. ChemKED-formatted files contain all of the information needed to simulate experimental
    data points, including the uncertainty of the data. ChemKED is based on the YAML data
    serialization language, and is intended as a human- and machine-readable standard for easy
    creation and automated use. Development of ChemKED and PyKED occurs openly on GitHub under the
    BSD 3-clause license, and contributions from the community are welcome. Plans for future
    development include support for experimental data from laminar flame, jet stirred reactor, and
    speciation measurements.
    %
\end{abstract}
